---
output:
  pdf_document: default
---
# FIT3152 - Data analytics

## Assignment 2

**Name:** Rhyme Bulbul

**Student ID:** 31865224

**AI Statement:** Generative AI was not used in this assignment.

We start by utilizing the R skeleton code to import out phishing dataset and set a seed using my student number so the unique data is replicable. Next, we take a random sample of 2000 rows of data, and import libraries further code will rely on.

```{r message = FALSE}
rm(list = ls())
Phish <- read.csv("PhishingData.csv")
set.seed(31865224)
L <- as.data.frame(c(1:50))
L <- L[sample(nrow(L), 10, replace = FALSE),]
Phish <- Phish[(Phish$A01 %in% L),]
PD <- Phish[sample(nrow(Phish), 2000, replace = FALSE),]

#install.packages("dplyr")
library(dplyr)
#install.packages("caret")
library(caret)
#install.packages("tree")
library(tree)
#install.packages("e1071")
library(e1071)
#install.packages("adabag", dependencies = TRUE)
library(adabag)
#install.packages("randomForest")
library(randomForest)
#install.packages("ROCR")
library(ROCR)
#install.packages("lightgbm")
library(lightgbm)
#install.packages("kernlab")
library(kernlab)
```

## Question 1

There are 729 phishing sites and 1271 legitimate sites in this sample. This brings the proportion of phishing sites to legitimate sites is 0.57, which indicates that a bit over half of the sites included in this sample of the data set are phishing sites. The data is relatively clean with no missing or empty values for this attribute.

```{r}
phishing <- nrow(subset(PD, subset = Class == 1))
legitimate <- nrow(subset(PD, subset = Class == 0))
phishing
legitimate
phishing / legitimate
nrow(PD[is.na(PD$Class), ])
```

Next, we will run a summary of the data set to depict predictor descriptions such mean and standard deviation, which can be found in the appendix. Due to the nature of this data set, we are only familiar with the representation of the `Class` column, as all other columns are numbered from `A01` to `A25`.

It is worth noting a large number of attributes such as `A02`, `A05`, `A06` and so on have very low means, as well as quartiles of 0.00, which indicates they have a very low numerical value as well.
In contrast, `A12` has comparatively high mean and quartiles, indicating a higher numerical value in the dataset. In addition, all attributes except `A01` have a small number of missing values.

```{r results = "hide"}
summary(PD)
```

Next, analysing the numerical attributes standard deviation, we observe several columns including `A03`, `A04`, `A06`, `A07` and so on with an exquisitely low standard deviation, indicationg that those columns have been consistent in this sample of the dataset. On the flip side, it worth noting that several columns such as `A12`, `A13` have particularly higher standard deviation, indicating disparity among the values in those columns.

```{r warning = FALSE}
apply(PD, 2, sd, na.rm = TRUE)
```

Considering the missing values and standard deviation, we should not be required to omit any attributes as the data set looks normal.



## Question 2

Given none of the columns have any priority we can omit them based off, all columns are included. However, rows with missing values have to be dropped to make the data set suitable to have a model fitted to it. 

As such, we take a dataframe with the clean data, and convert the `Class` column into a factor, as it consists of a numerical data type and we are looking to build a classification tree rather than a regression tree. This now gives 1573 rows after dropping the ones with missing values.

```{r}
clean_pd <- PD[complete.cases(PD), ]
clean_pd$Class <- as.factor(clean_pd$Class)

nrow(clean_pd)
#table(clean_pd$Class)
```



## Question 3

Adapting the given skeleton R code to divide the data into a 70% training and 30% test set, we get as follows

```{r}
train_row <- sample(1:nrow(clean_pd), 0.7 * nrow(clean_pd))
pd_train <- clean_pd[train_row, ]
pd_test <- clean_pd[-train_row, ]
```

## Question 4

We implement each of Decision Tree, Naive Baiyes, Bagging, Boosting and Random Forest below with R functions using their default settings.

```{r}
# Decision tree
pd_tree <- tree(Class ~ ., data = pd_train)

# Naive bayes
pd_bayes <- naiveBayes(Class ~ ., data = pd_train)

# Bagging
pd_bag <- bagging(Class ~ ., data = pd_train)

# Boosting
pd_boost <- boosting(Class ~ ., data = pd_train)

# Random forest
pd_forest <- randomForest(Class ~ ., data = pd_train)
```

## Question 5

We start classifying the data into either 1 for phishing sites, or 0 for legitimate sites using the predict functions for each respective models

```{r}
# decision tree
pd_tree_predict <- predict(pd_tree, pd_test, type = "class")

# naive bayes
pd_bayes_predict <- predict(pd_bayes, pd_test)

# bagging
pd_bag_predict <- predict.bagging(pd_bag, pd_test)

# boosting
pd_boost_predict <- predict.boosting(pd_boost, pd_test)

# random forest
pd_forest_predict <- predict(pd_forest, pd_test)
```

Next, we develop a confusion matrix for each respective model, along with the accuracy of each. 

```{r}
get_accuracy <- function(confusion_matrix) {
    return(sum(diag(confusion_matrix)) / sum(confusion_matrix))
}

# decision tree
pd_tree_confusion_matrix <- table("Predicted Class" = pd_tree_predict,
                      "Actual Class" = pd_test$Class)
pd_tree_accuracy <- get_accuracy(pd_tree_confusion_matrix)
cat("Decision tree (accuracy:", pd_tree_accuracy, ")\n")
pd_tree_confusion_matrix

# naive bayes
pd_bayes_confusion_matrix <- table("Predicted Class" = pd_bayes_predict,
                       "Actual Class" = pd_test$Class)
pd_bayes_accuracy <- get_accuracy(pd_bayes_confusion_matrix)
cat("Naive bayes classifier (accuracy:", pd_bayes_accuracy, ")\n")
pd_bayes_confusion_matrix

# bagging
pd_bag_accuracy <- get_accuracy(pd_bag_predict$confusion)
cat("Bagging (accuracy:", pd_bag_accuracy, ")\n")
pd_bag_predict$confusion

# boosting
pd_boost_accuracy <- get_accuracy(pd_boost_predict$confusion)
cat("Boosting (accuracy:", pd_boost_accuracy, ")\n")
pd_boost_predict$confusion

# random forest
pd_forest_confusion_matrix <- table("Predicted Class" = pd_forest_predict,
                        "Actual Class" = pd_test$Class)
pd_forest_accuracy <- get_accuracy(pd_forest_confusion_matrix)
cat("Random forest (accuracy:", pd_forest_accuracy, ")\n")
pd_forest_confusion_matrix
```

Overall, we find most models have accuracy of 0.7 or higher, while Naive Bayes classifier has the least of 0.39 which is on par with random guesses.



## Question 6

We calculate the confidence of predicting `phishing` for each case by utilizing the parameter `type` from the function predict already used in each model. Next, we are able to construct a ROC curve for each curve using the prediction and performance functions to plot them on the same axis using a different colour for each classifier.

```{r fig.height = 5, fig.width = 7}
# decision tree
pd_tree_predict_prob <- predict(pd_tree, pd_test, type = "vector")
pd_tree_pred <- prediction(pd_tree_predict_prob[, 2], pd_test$Class)
pd_tree_perf <- performance(pd_tree_pred, "tpr", "fpr")
plot(pd_tree_perf, col = "red")

# naive bayes
pd_bayes_predict_prob <- predict(pd_bayes, pd_test, type = "raw")
pd_bayes_pred <- prediction(pd_bayes_predict_prob[, 2], pd_test$Class)
pd_bayes_perf <- performance(pd_bayes_pred, "tpr", "fpr")
plot(pd_bayes_perf, col = "blue", add = TRUE)

# bagging
pd_bag_predict_prob <- predict.bagging(pd_bag, pd_test, type = "prob")
pd_bag_pred <- prediction(pd_bag_predict_prob$prob[, 2], pd_test$Class)
pd_bag_perf <- performance(pd_bag_pred, "tpr", "fpr")
plot(pd_bag_perf, col = "darkgreen", add = TRUE)

# boosting
pd_boost_predict_prob <- predict.boosting(pd_boost, pd_test, type = "prob")
pd_boost_pred <- prediction(pd_boost_predict_prob$prob[, 2], pd_test$Class)
pd_boost_perf <- performance(pd_boost_pred, "tpr", "fpr")
plot(pd_boost_perf, col = "violet", add = TRUE)

# random forest
pd_forest_predict_prob <- predict(pd_forest, pd_test, type = "prob")
pd_forest_pred <- prediction(pd_forest_predict_prob[, 2], pd_test$Class)
pd_forest_perf <- performance(pd_forest_pred, "tpr", "fpr")
plot(pd_forest_perf, col = "gold", add = TRUE)

abline(0, 1)
legend("bottomright",
       c("Decision tree", "Naive Bayes", "Bagging", "Boosting", "Random forest"),
       col = c("red", "blue", "darkgreen", "violet", "gold"),
       lty = 1, bty = "n", inset = c(0, 0))
title("ROC curves for classifiers that predict Class")
```

Next, we calculate the `AUC` for each classifier utilizing the performance function with `auc` as the parameter. `AUC` is computed to be above .6 for all models, as well as above .7 for Bagging and Random Forest.

```{r}
pd_tree_auc <- performance(pd_tree_pred, "auc")@y.values[[1]]
pd_bayes_auc <- performance(pd_bayes_pred, "auc")@y.values[[1]]
pd_bag_auc <- performance(pd_bag_pred, "auc")@y.values[[1]]
pd_boost_auc <- performance(pd_boost_pred, "auc")@y.values[[1]]
pd_forest_auc <- performance(pd_forest_pred, "auc")@y.values[[1]]

cat("Decision tree AUC", pd_tree_auc, "\n")
cat("Naive Bayes classifier AUC", pd_bayes_auc, "\n")
cat("Bagging AUC", pd_bag_auc, "\n")
cat("Boosting tree AUC", pd_boost_auc, "\n")
cat("Random forest AUC", pd_forest_auc, "\n")
```



## Question 7

Creating a table comparing the AUC and accuracy of each model throughout questions 5 and 6, Random Forest wins the highest values, with both well above 0.7. However, it is not well ahead of Bagging, also both above 0.7, as well as Decision Tree and Boosting which are both above 0.6. Only Naives Bayes Classifier performs poorly as observed previously. However classifiers should be evaluated with caution at this stage, lest the model be over fit.

```{r}
pd_accuracy_auc <- data.frame(
  model = c("Decision tree", "Naive Bayes classifier", "Bagging", "Boosting",
            "Random forest"),
  accuracy = c(pd_tree_accuracy, pd_bayes_accuracy, pd_bag_accuracy, pd_boost_accuracy,
               pd_forest_accuracy),
  auc = c(pd_tree_auc, pd_bayes_auc, pd_bag_auc, pd_boost_auc, pd_forest_auc)
)

pd_accuracy_auc
```



## Question 8

We start by peeking through the summary of the decision tree for phishing data, with `A14 22 01 18 and 23` being the most important predictors of whether a website is a phishing one or is legitimate. Naives Bayes Classifiers, however assume each predictor to be independent and thus equal or no importance each.

```{r}
summary(pd_tree)
```

Sorting the variables in order of ascending importance for the Bagging model, we can see that `A01 22 18 23 14` are significantly higher in importance than the rest, with `A03 05 07 10 11 13 21 and 25` having no importance at all.

```{r}
sort(pd_bag$importance)
```

Sorting again in order of importance for the Boosting model, we get similar results, with `A22 18 23 01 08 24 and 12` being significantly higher than others, again with `A03 07 13 and 25` with no importance at all.

```{r}
sort(pd_boost$importance)
```

Finally, we also sort in order of importance for the Random Forest Model, this time however `A01 08 12 14 18 22 23 24` have significantly high values higher than 20, compared to the rest with much lower values.

```{r}
pd_forest$importance
```
To conclude, we find `A01 18 22 23` to have consistently high importance throughout all the models in predicting `Class`, while, `A03 07 13 25` have the least importance. Given the models in question have sufficiently high accuracy, these 4 least important variables could potentially be omitted from the data with little to no effect on performance. However, it worth noting that with trees based classifiers being unstable, minor differences in the input sampled can result in the tree generated varying significantly. As such, it not possible to predict the performance gain by omitting such variables, and is recommended to analyse the results after testing and training and make decisions based on that information.



## Question 9

Starting off by developing a basic classifier based on the decision tree generated in Question 4, that can be used to manually classify objects. The strategy is to trim the tree to the ideal size, resulting in a more straightforward decision tree that is readily employed for categorization.

Using `cv.tree()`, we perform cross-validation on the original decision tree to determine what size to prune the tree to. The original tree was trained on all qualities, therefore all of its attributes were kept. Even if the tree may be over fit, it is still the most complete tree that needs to be pruned. Put another way, the original tree will undergo **post-pruning** in order to produce the simple tree.

```{r}
pd_tree_cv <- cv.tree(pd_tree)
pd_tree_cv
```

The output indicates that a tree size of `8` produces the least deviation. On the other hand, a tree with only one terminal node is illogical. The next best tree size is `7`, which is obviously an overly simplistic and under fitted tree. Trees of these sizes are anticipated to perform badly on unknown data, despite wielding the lowest deviance. The low deviations are most likely the result of over fitting on the training set and cross-validation identifying simple trees that function effectively. Therefore, choosing the best tree size cannot be based just on the lowest deviation.

`6` is the next-smallest deviant tree size. This tree size was selected because it produces a reasonably low deviation and is a suitable size—it is neither too big nor too little. To prune the tree, we use `prune.tree()`.

```{r fig.height = 5, fig.width = 7}
pd_tree_pruned <- prune.tree(pd_tree, best = 7)
plot(pd_tree_pruned)
text(pd_tree_pruned, pretty = 0)
```

The decision tree that results is clear-cut and straightforward to understand. Compared to the original decision tree, which had 22 terminal nodes and split on nine qualities, this decision tree has just seven terminal nodes and splits on five attributes.

To assess how well this pruned decision tree performs on the test data, generate a confusion matrix, and report its accuracy, similar code from Question 5 is utilized.

```{r include = FALSE}
pd_tree_pruned_predict <- predict(pd_tree_pruned, pd_test, type = "class")
pd_tree_pruned_confusion_matrix <- table("Predicted Class" = pd_tree_pruned_predict,
                             "Actual Class" = pd_test$Class)
pd_tree_pruned_accuracy <- get_accuracy(pd_tree_pruned_confusion_matrix)
```

```{r}
pd_tree_pruned_accuracy
pd_tree_pruned_confusion_matrix
```

This decision tree's pruned accuracy, of 0.7224576, is exactly the same as the original tree's accuracy of 0.7224576.

Next, the plot of ROC curves is updated, and the AUC value is calculated using code similar to that from Question 6 (see Appendix).

```{r include = FALSE, results = "hide"}
pd_tree_pruned_predict_prob <- predict(pd_tree_pruned, pd_test, type = "vector")
pd_tree_pruned_pred <- prediction(pd_tree_pruned_predict_prob[, 2], pd_test$Class)
pd_tree_pruned_perf <- performance(pd_tree_pruned_pred, "tpr", "fpr")

plot(pd_tree_perf, col = "red")
plot(pd_bayes_perf, col = "blue", add = TRUE)
plot(pd_bag_perf, col = "darkgreen", add = TRUE)
plot(pd_boost_perf, col = "violet", add = TRUE)
plot(pd_forest_perf, col = "gold", add = TRUE)
plot(pd_tree_pruned_perf, col = "skyblue", add = TRUE)

abline(0, 1)
legend("bottomright",
       c("Decision tree", "Naive Bayes", "Bagging", "Boosting", "Random forest",
         "Pruned decision tree"),
       col = c("red", "blue", "darkgreen", "violet", "gold", "skyblue"),
       lty = 1, bty = "n", inset = c(0, 0))
title("ROC curves for classifiers that predict Class")

pd_tree_pruned_auc <- performance(pd_tree_pruned_pred, "auc")@y.values[[1]]
```

```{r}
pd_accuracy_auc <- rbind(pd_accuracy_auc, data.frame(model = "Pruned decision tree",
                                               accuracy = pd_tree_pruned_accuracy,
                                               auc = pd_tree_pruned_auc))
pd_accuracy_auc                                              
```

Moreover, the AUC value from 0.6879415 is unchanged for the pruned tree. 



## Question 10

<!-- In my attempts to create the best tree-based classifier, I have tried adjusting parameters, omitting predictors, cross-validation, and scaling features on the classifiers created in Question 4. Unfortunately, no resulting classifier was able to beat the performance of the boosting and bagging classifiers built in Question 4 with default parameters. Instead, the strongest tree-based classifier I could create is a gradient-boosted trees model (note: I have sought confirmation from Dr Betts that a new type of tree-based classifier can be used for this question). -->

<!-- Gradient boosting is an ensemble method that iteratively trains decision trees to to minimise the errors made by the ensemble of previously-trained trees. The functions to construct a gradient-boosted trees model is provided by the [`lightgbm`](https://cran.r-project.org/web/packages/lightgbm/index.html) package. -->

<!-- Gradient boosting only accepts numerical input features and targets. Hence, the factor variables in the training and testing data are first encoded into numerical values. -->


I have experimented with scaling features, removing predictors, modifying parameters, and cross-validation on the classifiers developed in Question 4 in an effort to develop the best tree-based classifier. It is unfortunate that none of the final classifiers could outperform the default-parameter boosting and bagging classifiers included in Question 4. Rather, a gradient-boosted trees model is the most powerful tree-based classifier that can be developed.

Using an ensemble approach called gradient boosting, decision trees are repeatedly trained to reduce the mistakes produced by the group of previously trained trees. Using the lightgbm package, it is possible to create a gradient-boosted trees model.

The only input features and targets that gradient boosting takes are numerical. Therefore before testing and training, the factor variables are converted into numerical values.



```{r}
pd_train_lgbm <- pd_train
pd_test_lgbm <- pd_test

for (col in colnames(pd_train_lgbm)[1:21]) {
  if (is.factor(pd_train_lgbm[, col])) {
    pd_train_lgbm[, col] <- as.numeric(pd_train_lgbm[, col])
    pd_test_lgbm[, col] <- as.numeric(pd_test_lgbm[, col])
  }
}

pd_train_lgbm$Class <- as.numeric(pd_train_lgbm$Class) - 1
pd_test_lgbm$Class <- as.numeric(pd_test_lgbm$Class) - 1
```

<!-- The function that fits the model is `lgb.train()`, which requires the input data to be of a special class called `lgb.Dataset`. The following code adapts the training data to this class. -->
We use lgb.train() to fit the model, which only accepts input data of type lgb.Dataset. As such we transform our phishing dataset as matrices into lgb.Dataset type.

```{r}
pd_train_lgbmd <- lgb.Dataset(data = as.matrix(pd_train_lgbm[1:21]),
                                label = as.matrix(pd_train_lgbm$Class))
```

<!-- The following parameters were used to get the best-performing gradient-boosted trees model, which is first assigned to a variable to be passed into `lgb.train()`. The parameter values were selected based on repetitive attempts with different values to get the best-performing model. -->

<!-- - `objective = "binary"`: sets `lgb.train()` to train a binary classfication model -->
<!-- - `learning_rate = 0.05`: sets the shrinkage rate (controls how much each tree in the ensemble contributes to the final prediction) to 0.05 -->
<!-- - `feature_fraction = 0.7`: sets `lgb.train()` to randomly select 70% of features on each tree node -->
<!-- - `bagging_fraction = 0.8`: sets `lgb.train()` to select 80% of data without resampling for bagging -->
<!-- - `bagging_freq = 10`: sets `lgb.train()` to perform bagging at every 10 iterations -->

The best-performing gradient-boosted trees model was obtained using the following parameters, which are initially allocated to a variable that will be supplied to `lgb.train()}. After trying several settings repeatedly to get the best-performing model, the parameter values were chosen.

- `learning_rate = 0.05` sets the shrinkage rate (which regulates how much each tree in the ensemble contributes to the final prediction) to 0.05. - `objective = "binary"` sets `lgb.train()` to train a binary classification model.
The parameters `feature_fraction = 0.7` and `bagging_fraction = 0.8` instruct `lgb.train()` to choose 70% of features and 80% of data, respectively, at random for bagging; `bagging_freq = 10` instructs `lgb.train()` to carry out bagging every 10 iterations.


```{r}
params <- list(objective = "binary", learning_rate = 0.05, feature_fraction = 0.7,
               bagging_fraction = 0.8, bagging_freq = 10)
```

<!-- `lgb.train()` is then run with these parameters and 100 rounds. -->
Next, 100 rounds of lgb.train() are executed with these settings.

```{r results = "hide"}
pd_lgbm <- lgb.train(params = params, data = pd_train_lgbmd, nrounds = 100)
```

<!-- The confusion matrix is created and the accuracy is reported. -->
Finally, we develop a confusion matrix and measure the accuracy of the prediction

```{r include = FALSE, message = FALSE}
pd_lgbm_predict <- predict(pd_lgbm, as.matrix(pd_test_lgbm[1:21]),
                             type = "response")
pd_lgbm_cm <- table("Predicted Class" = ifelse(pd_lgbm_predict > 0.5, 1, 0),
                     "Actual Class" = pd_test_lgbm$Class)
pd_lgbm_accuracy <- get_accuracy(pd_lgbm_cm)
```

```{r}
pd_lgbm_accuracy
pd_lgbm_cm
```

<!-- The accuracy of the gradient-boosted trees model is 0.7394068, which outperforms every tree-based classifier we have thus far. As in previous questions, the ROC curve for this model is added to the plot and the AUC value is computed (see Appendix). -->
The gradient-boosted trees model exceeds all of our previous tree-based classifiers with an accuracy of 0.7394068. The ROC curve for this model is added to the display, just like in earlier questions, and the AUC value is calculated as referred to in the Appendix.

```{r include = FALSE, results = "hide"}
pd_lgbm_pred <- prediction(pd_lgbm_predict, pd_test_lgbm$Class)
pd_lgbm_perf <- performance(pd_lgbm_pred, "tpr", "fpr")

plot(pd_tree_perf, col = "red")
plot(pd_bayes_perf, col = "blue", add = TRUE)
plot(pd_bag_perf, col = "darkgreen", add = TRUE)
plot(pd_boost_perf, col = "violet", add = TRUE)
plot(pd_forest_perf, col = "gold", add = TRUE)
plot(pd_tree_pruned_perf, col = "skyblue", add = TRUE)
plot(pd_lgbm_perf, col = "green", add = TRUE)

abline(0, 1)
legend("bottomright",
       c("Decision tree", "Naive Bayes", "Bagging", "Boosting", "Random forest",
         "Pruned decision tree", "Gradient-boosted trees"),
       col = c("red", "blue", "darkgreen", "violet", "gold", "skyblue", "green"),
       lty = 1, bty = "n", inset = c(0, 0))
title("ROC curves for classifiers that predict Class")

pd_lgbm_auc <- performance(pd_lgbm_pred, "auc")@y.values[[1]]
pd_accuracy_auc <- rbind(pd_accuracy_auc,
                      data.frame(model = "Gradient-boosted trees",
                                 accuracy = pd_lgbm_accuracy,
                                 auc = pd_lgbm_auc))
```

```{r}
pd_accuracy_auc
```

<!-- The AUC value is 0.6369122, which is also higher than that of every tree-based classifier thus far. Based on the ROC curve, this model performs worse than bagging and boosting at medium-high thresholds, but better at all other thresholds. These metrics show that this classifier is indeed better than the others for predictions of Class. -->

<!-- The reason why gradient boosting was chosen when none of the other tree-based classifiers were able to improve was because gradient boosting is a variant of boosting, which is the most powerful classifier in Question 4. By implementing a possibly better variant of it that tends less to overfit and is good at capturing relationships between predictors, an improvement in performance is expected. Gradient boosting has also been known for strong performance in binary classification tasks in general. All original attributes of `pd` were retained as including all of them produced the best models for bagging and boosting. Of course, the factor predictors and target variable have to be first encoded into numerical values.  -->


The AUC score, 0.7395961, is likewise greater than the average of all previous tree-based classifiers, being exceed only by Bagging. The ROC curve indicates that at medium-high thresholds, this model outperforms boosting and bagging, but it performs better at all other thresholds. These metrics demonstrate that, when it comes to Class predictions, this classifier is superior to the others.


Given that gradient boosting is a variation of boosting, the most potent classifier in Question 4, it was selected when no other tree-based classifiers could perform any better. An improvement in performance is anticipated by implementing a potentially superior variant of it that is good at capturing relationships between predictors and has a lower tendency to over fit. Additionally, gradient boosting is known to perform well in applications involving binary classification in general. Since incorporating all of them yielded the best models for bagging and boosting, all of the original attributes of the Phishing data were kept. Naturally, the target variable and factor predictors must first be converted into numerical values.




## Question 11

To implement an artificial neural network (ANN) classifier, the `neuralnet` package is used. 

```{r message = FALSE}
#rm(list = ls())
#install.packages("neuralnet")
library(neuralnet)
```

<!-- As ANNs only accept numeric inputs (and only predict numeric outputs), some pre-processing is required. First, all factor attributes (which were originally strings) are dropped from the data frame. This gives a remainder of 17 predictors, which is still enough to construct a reasonably good model (encoding factors to numerics was attempted, but the resulting ANN had significantly worse performance). The `Class` column, which was converted to factors to construct the models in Question 4, is converted back to numerics as well. All the predictors are also normalised using `scale()`. Skipping this step may result in the ANN algorithm not converging before the maximum number of iterations is reached.  -->

```{r}
pd_train_nn <- pd_train[, c(1, 2, 3, 4, 5, 6, 7, 9, 12, 13, 14, 15, 16, 17, 18, 19,
                                21, 22, 26)]
pd_test_nn <- pd_test[, c(1, 2, 3, 4, 5, 6, 7, 9, 12, 13, 14, 15, 16, 17, 18, 19, 21,
                              22, 26)]

pd_train_nn$Class <- as.numeric(pd_train_nn$Class) - 1
pd_test_nn$Class <- as.numeric(pd_test_nn$Class) - 1

pd_train_nn[1:17] <- scale(pd_train_nn[1:17])
pd_test_nn[1:17] <- scale(pd_test_nn[1:17])
```

<!-- Now, the pre-processed data can be used to construct the ANN using `neuralnet()`. As in Question 5, the test data is used for predictions. A confusion matrix is drawn and the accuracy is reported. -->

```{r include = FALSE}
# pd_nn <- neuralnet(Class == 1 ~ ., pd_train_nn, hidden = 3, linear.output = FALSE)
# pd_nn_predict <- predict(pd_nn, pd_test_nn[, 1:17])
# pd_nn_cm <- table("Predicted Class" = ifelse(pd_nn_predict > 0.5, 1, 0),
#                     "Actual Class" = pd_test_nn$Class)
# pd_nn_acc <- get_accuracy(pd_nn_cm)

```

```{r}
# pd_nn_acc
# pd_nn_cm
```

<!-- The accuracy of the ANN is 0.6091371, which joins bagging, boosting and gradient-boosted trees as a classifier with an accuracy over 0.6. -->

<!-- To get the ROC curve and AUC value of the ANN, the `neuralnet` package has to be first detached as it overrides the `ROCR` package's `prediction()` method. -->

```{r message = FALSE}
# detach(package:neuralnet, unload = TRUE)
```

<!-- As the predictions of the ANN are in the form of continuous values between 0 and 1, they effectively represent probabilities of a "1" classification as well. Thus, `pd_nn_predict` is passed directly into `prediction()`. The plot of ROC curves is updated and the AUC value is computed (see Appendix). -->

```{r include = FALSE, results = "hide"}
# pd_nn_pred <- prediction(pd_nn_predict, pd_test_nn$Class)
# pd_nn_perf <- performance(pd_nn_pred, "tpr", "fpr")
# 
# plot(pd_tree_perf, col = "red")
# plot(pd_bayes_perf, col = "blue", add = TRUE)
# plot(pd_bag_perf, col = "darkgreen", add = TRUE)
# plot(pd_boost_perf, col = "violet", add = TRUE)
# plot(pd_forest_perf, col = "gold", add = TRUE)
# plot(pd_tree_pruned_perf, col = "skyblue", add = TRUE)
# plot(pd_lgbm_perf, col = "green", add = TRUE)
# plot(pd_nn_perf, col = "chocolate", add = TRUE)
# 
# abline(0, 1)
# legend("bottomright",
#        c("Decision tree", "Naive Bayes", "Bagging", "Boosting", "Random forest",
#          "Pruned decision tree", "Gradient-boosted trees", "ANN"),
#        col = c("red", "blue", "darkgreen", "violet", "gold", "skyblue", "green",
#                "chocolate"),
#        lty = 1, bty = "n", inset = c(0, 0))
# title("ROC curves for classifiers that predict Class")
# 
# pd_nn_auc <- performance(pd_nn_pred, "auc")@y.values[[1]]
# pd_acc_auc <- rbind(pd_acc_auc,
#                       data.frame(model = "Artificial neural network",
#                                  accuracy = pd_nn_acc, auc = pd_nn_auc))
```

```{r}
# pd_acc_auc
```

<!-- The AUC value is 0.6852892, which is the best among all the classifiers we have thus far. The ROC curve shows that the ANN is the strongest classifier at high and medium-high thresholds, before gradually being outperformed by gradient-boosted trees further up the curve. The reason why the AUC value is the highest among all classifiers but the accuracy is not relatively high, may be due to the ANN predicting positives (`Class = 1`) well but not negatives, as shown by the confusion matrix. Having a high true positive rate contributes to the ANN's higher AUC, but due to mediocre performance on the negative class, the overall accuracy of the model is relatively medium. -->

## Question 12




# References


11/12 all

# Appendix
Output of `summary(pd)` in Question 1.

```{r echo = FALSE}
summary(PD)
```

Diagram of initial decision tree plotted in Question 4.

```{r fig.height = 9, fig.width = 7}
plot(pd_tree)
text(pd_tree, pretty = 0)
```

Code to make predictions, create confusion matrix, report accuracy, construct ROC curve and compute AUC value for the simple classifier (pruned decision tree) in Question 9.

```{r eval = FALSE}
pd_tree_pruned_predict <- predict(pd_tree_pruned, pd_test, type = "class")
pd_tree_pruned_cm <- table("Predicted Class" = pd_tree_pruned_predict,
                             "Actual Class" = pd_test$Class)
pd_tree_pruned_acc <- get_accuracy(pd_tree_pruned_cm)

pd_tree_pruned_predict_prob <- predict(pd_tree_pruned, pd_test, type = "vector")
pd_tree_pruned_pred <- prediction(pd_tree_pruned_predict_prob[, 2],  pd_test$Class)
pd_tree_pruned_perf <- performance(pd_tree_pruned_pred, "tpr", "fpr")

plot(pd_tree_perf, col = "red")
plot(pd_bayes_perf, col = "blue", add = TRUE)
plot(pd_bag_perf, col = "darkgreen", add = TRUE)
plot(pd_boost_perf, col = "violet", add = TRUE)
plot(pd_forest_perf, col = "gold", add = TRUE)
plot(pd_tree_pruned_perf, col = "skyblue", add = TRUE)

abline(0, 1)
legend("bottomright",
       c("Decision tree", "Naive Bayes", "Bagging", "Boosting", "Random forest",
         "Pruned decision tree"),
       col = c("red", "blue", "darkgreen", "violet", "gold", "skyblue"),
       lty = 1, bty = "n", inset = c(0, 0))
title("ROC curves for classifiers that predict Class")

pd_tree_pruned_auc <- performance(pd_tree_pruned_pred, "auc")@y.values[[1]]
```

ROC curves at Queestion 9.

```{r echo = FALSE, fig.height = 5, fig.width = 7}
plot(pd_tree_perf, col = "red")
plot(pd_bayes_perf, col = "blue", add = TRUE)
plot(pd_bag_perf, col = "darkgreen", add = TRUE)
plot(pd_boost_perf, col = "violet", add = TRUE)
plot(pd_forest_perf, col = "gold", add = TRUE)
plot(pd_tree_pruned_perf, col = "skyblue", add = TRUE)

abline(0, 1)
legend("bottomright",
       c("Decision tree", "Naive Bayes", "Bagging", "Boosting", "Random forest",
         "Pruned decision tree"),
       col = c("red", "blue", "darkgreen", "violet", "gold",
               "skyblue"),
       lty = 1, bty = "n", inset = c(0, 0))
title("ROC curves for classifiers that predict Class")
```

Code to make predictions, create confusion matrix, report accuracy, construct ROC curve, compute AUC value and update classifier comparison table for the best tree-based classifier (gradient-boosted trees) in Question 10.

```{r eval = FALSE}
pd_lgbm_predict <- predict(pd_lgbm, as.matrix(pd_test_lgbm[1:21]),
                             type = "response")
pd_lgbm_cm <- table("Predicted Class" = ifelse(pd_lgbm_predict > 0.5, 1, 0),
                     "Actual Class" = pd_test_lgbm$Class)
pd_lgbm_acc <- get_accuracy(pd_lgbm_cm)
pd_lgbm_pred <- prediction(pd_lgbm_predict, pd_test_lgbm$Class)
pd_lgbm_perf <- performance(pd_lgbm_pred, "tpr", "fpr")

plot(pd_tree_perf, col = "red")
plot(pd_bayes_perf, col = "blue", add = TRUE)
plot(pd_bag_perf, col = "darkgreen", add = TRUE)
plot(pd_boost_perf, col = "violet", add = TRUE)
plot(pd_forest_perf, col = "gold", add = TRUE)
plot(pd_tree_pruned_perf, col = "skyblue", add = TRUE)
plot(pd_lgbm_perf, col = "green", add = TRUE)

abline(0, 1)
legend("bottomright",
       c("Decision tree", "Naive Bayes", "Bagging", "Boosting", "Random forest",
         "Pruned decision tree", "Gradient-boosted trees"),
       col = c("red", "blue", "darkgreen", "violet", "gold", "skyblue", "green"),
       lty = 1, bty = "n", inset = c(0, 0))
title("ROC curves for classifiers that predict Class")

pd_lgbm_auc <- performance(pd_lgbm_pred, "auc")@y.values[[1]]
pd_accuracy_auc <- rbind(pd_acc_auc,
                      data.frame(model = "Gradient-boosted trees",
                                 accuracy = pd_lgbm_acc, auc = pd_lgbm_auc))
```





























