# FIT3152 - Data analytics

## Assignment 2

**Name:** Rhyme Bulbul

**Student ID:** 31865224

**AI Statement:** Generative AI was not used in this assignment.

<!-- - The main body of this report is almost 15 pages. This length is due to some long code blocks and the sizes of the decision tree diagram and ROC curve plots that have to be big to make them readable. All other pages are the Appendix, which include repeated code and outputs. -->
<!-- - Due to length, from Question 9 onwards, similar code from Questions 4 and 5 used to make predictions, create confusion matrices, report accuracies, construct ROC curves and compute AUC values will not be shown in the main report but in the Appendix. ROC curve plots will also not be shown in the main report except that of Question 12, which shows the curves for all classifiers. -->

<!-- My unique dataset of 2000 rows is first created with the following code from the specifications. Libraries needed to perform the tasks are imported as well. -->

We start by utilizing the R skeleton code to import out phishing dataset and set a seed using my student number so the unique data is replicable. Next, we take a random sample of 2000 rows of data, and import libraries further code will rely on.

```{r message = FALSE}
rm(list = ls())
Phish <- read.csv("PhishingData.csv")
set.seed(31865224)
L <- as.data.frame(c(1:50))
L <- L[sample(nrow(L), 10, replace = FALSE),]
Phish <- Phish[(Phish$A01 %in% L),]
PD <- Phish[sample(nrow(Phish), 2000, replace = FALSE),]

library(dplyr)
library(caret)
library(tree)
library(e1071)
#install.packages("adabag")
#library(adabag)
#install.packages("randomForest")
#library(randomForest)
library(ROCR)
library(lightgbm)
#install.packages("kernlab")
#library(kernlab)
```

## Question 1

<!-- The proportion of days when it is more humid than the previous day, and the opposite, can be computed by subsetting the data based on the value of `MHT`, which indicates whether the day in the next entry is more humid than that of the current one. -->

There are 729 phishing sites and 1271 legitimate sites in this sample. This brings the proportion of phishing sites to legitimate sites is 0.57, which indicates that a bit over half of the sites included in this sample of the data set are phishing sites. The data is relatively clean with no missing or empty values for this attribute.

```{r}
phishing <- nrow(subset(PD, subset = Class == 1))
legitimate <- nrow(subset(PD, subset = Class == 0))
phishing
legitimate
phishing / legitimate
nrow(PD[is.na(PD$Class), ])
```

<!-- Based on the results, the proportion is about 1.03, which means their respective frequencies are close to equal. The `MHT` field is also empty for 109 entries, so the true proportion may differ slightly as roughly 5% of the data has no target variable value. -->

<!-- `summary()` is then run to get descriptions of the predictors. Due to its length, the output is shown in the Appendix. -->
Next, we will run a summary of the data set to depict predictor descriptions such mean and standard deviation, which can be found in the appendix. Due to the nature of this data set, we are only familiar with the representation of the `Class` column, as all other columns are numbered from `A01` to `A25`.

It is worth noting a large number of attributes including `AXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX` have very low means, as well as quartiles of 0.00, which indicates they have a very low numerical value as well.
In contrast, `A12` has comparatively high mean and quartiles, indicating a higher numerical value in the dataset. In addition, all attributes except `A01` have a small number of missing values.

```{r results = "hide"}
summary(PD)
```

<!-- Among the numerical real-valued attributes, most of the means are between 2 and 22. Some exceptions are `WindGustSpeed`, with a mean of 43.62, and more notably, `Pressure9am` and `Pressure3pm`, with means more than 1000 (which is normal for the scale at which atmospheric pressure is measured). Some interesting mean values are `Rainfall`'s mean of 2.941 despite having a max value of 156.4, and `WindGustSpeed`'s mean despite having a max value of 130. In both cases, this indicates a presence of a few extreme values. -->

<!-- Additionally, one noteworthy observation is that there are NA values for every predictor except `Location`, with some predictors having significantly more than the rest (eg. `Sunshine` has 842 NA's compared to `Year`'s 22). -->

<!-- Then, this code is run to compute the standard deviations of the numerical attributes. -->

Next, analysing the numerical attributes standard deviation, we observe several columns including `AXXXXXXXXXXXXXXXXXXXXX` with an exquisitely low standard deviation, indicationg that those columns have been consistent in this sample of the dataset. On the flip side, it worth noting that several columns such as `AXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXx` have particularly higher standard deviation, indicating disparity among the values in those columns.

```{r warning = FALSE}
apply(PD, 2, sd, na.rm = TRUE)
```

Considering the missing values and standard deviation, we should not be required to omit any attributes as the data set looks normal.
<!-- As expected, `WindGustSpeed` and `Rainfall` are the two real-valued attributes with the largest standard deviations due to their extreme values. The smallest standard deviation belongs to `Cloud3pm`, meaning that the fraction of sky obscured by cloud at 3pm has been consistent, for this dataset. -->

<!-- The only attributes that may, from first glance, be worth omitting are `Location` and `Year`, as a country's yearly climate is usually expected to be consistent. But due to Australia's size, diverse climate and geography, and climate change in recent years, both `Location` and `Year` might have a significant influence on `MHT` and should be kept. All other predictors are closely related to the weather and are thus assumed to have predictive power on `MHT` as well. -->


## Question 2

<!-- As all the predictors seem to have relatively equal importance based on their descriptions in the specifications (and it is generally a good idea to to build initial classifers using all attributes and skim them down from there), no columns are dropped. The only pre-processing required is to remove all rows with NA values to make the data suitable for model fitting. -->

<!-- The following code extracts only complete entries from `WAUS` into a new data frame called `waus_clean`, and `MHT` is also converted to factors as its data type is numerical, which would result in a regression tree being built instead of a classification tree that classifies data points into 0 and 1. -->

Given none of the columns have any priority we can omit them based off, all columns are included. However, rows with missing values have to be dropped to make the data set suitable to have a model fitted to it. 

As such, we take a dataframe with the clean data, and convert the `Class` column into a factor, as it consists of a numerical data type and we are looking to build a classification tree rather than a regression tree. This now gives 1573 rows after dropping the ones with missing values.

```{r}
clean_pd <- PD[complete.cases(PD), ]
clean_pd$Class <- as.factor(clean_pd$Class)

nrow(clean_pd)
#table(clean_pd$Class)
```

<!-- The outputs of the last two lines shows that the processed dataset now contains 655 rows, and the number of unique locations has been reduced to 5. -->

## Question 3
<!-- The adapted code to divide my data with a 70% to 30% train-test split is as follows. -->
Adapting the given skeleton R code to divide the data into a 70% training and 30% test set, we get as follows

```{r}
train_row <- sample(1:nrow(clean_pd), 0.7 * nrow(clean_pd))
pd_train <- clean_pd[train_row, ]
pd_test <- clean_pd[-train_row, ]
```

## Question 4

<!-- A classification model of each technique is constructed with default settings. -->
We implement each of Decision Tree, Naive Baiyes, Bagging, Boosting and Random Forest below with R functions using their default settings.

```{r}
# Decision tree
pd_tree <- tree(Class ~ ., data = pd_train)

# Naive bayes
pd_bayes <- naiveBayes(Class ~ ., data = pd_train)

# Bagging
#pd_bag <- bagging(Class ~ ., data = pd_train)

# Boosting
#pd_boost <- boosting(Class ~ ., data = pd_train)

# Random forest
#pd_forest <- randomForest(Class ~ ., data = pd_train)
```

## Question 5
`TODO!!!`
<!-- Each data point in the test data is classified into either `1` (more humid tomorrow) or `0` (less humid tomorrow) using `predict()` (`predict.bagging()` and `predict.boosting()` for bagging and boosting respectively). -->

```{r}
# decision tree
pd_tree_predict <- predict(pd_tree, pd_test, type = "class")

# naive bayes
pd_bayes_predict <- predict(pd_bayes, pd_test)

# bagging
#pd_bag_predict <- predict.bagging(pd_bag, pd_test)

# boosting
#pd_boost_predict <- predict.boosting(pd_boost, pd_test)

# random forest
#pd_forest_predict <- predict(pd_forest, pd_test)
```

<!-- A confusion matrix is then created for each model, displayed alongside each model's accuracy, which is the total number of accurate predictions divided by total number of observations. A function is used to compute the accuracy for each confusion matrix, using the code `sum(diag(cm)) / sum(cm)`, where `cm` is a given confusion matrix. -->

```{r}
get_accuracy <- function(cm) {
    return(sum(diag(cm)) / sum(cm))
}

# decision tree
pd_tree_cm <- table("Predicted Class" = pd_tree_predict,
                      "Actual Class" = pd_test$Class)
pd_tree_acc <- get_accuracy(pd_tree_cm)
cat("Decision tree (accuracy:", pd_tree_acc, ")\n")
pd_tree_cm

# naive bayes
pd_bayes_cm <- table("Predicted Class" = pd_bayes_predict,
                       "Actual Class" = pd_test$Class)
pd_bayes_acc <- get_accuracy(pd_bayes_cm)
cat("Naive bayes classifier (accuracy:", pd_bayes_acc, ")\n")
pd_bayes_cm

# bagging
# pd_bag_acc <- get_accuracy(pd_bag_predict$confusion)
# cat("Bagging (accuracy:", pd_bag_acc, ")\n")
# pd_bag_predict$confusion

# boosting
# pd_boost_acc <- get_accuracy(pd_boost_predict$confusion)
# cat("Boosting (accuracy:", pd_boost_acc, ")\n")
# pd_boost_predict$confusion

# random forest
# pd_forest_cm <- table("Predicted Class" = pd_forest_predict,
#                         "Actual Class" = pd_test$Class)
# pd_forest_acc <- get_accuracy(pd_forest_cm)
# cat("Random forest (accuracy:", pd_forest_acc, ")\n")
# pd_forest_cm
```

<!-- Among the classifiers, only bagging and boosting have an accuracy above 0.6. In fact, they have identical accuracies (0.6345178). The other three classifiers have accuracies closer to that of random guessing, with the naive Bayes classifier having the worst accuracy at 0.5482234. -->

## Question 6











































